\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2023-06-01 2-cubble.Rmd
section 2.4 compatibility with tsibble and sf

The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object. Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e. \code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e. \code{meteo_ts}. The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:

```{r echo = TRUE}
(res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
class(res)
class(res$ts[[1]])
```

@@

|Category | Details |  Functions|
|---|---|--------------|
|base R | subset | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
|tidyverse | dplyr |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `slice`, `group_by`, `ungroup`|
|tidyverse| tidyr| | 
|tidyverse| pillar| `type_sum` | 
|sf| | | 
|tsibble| | | 
|cubble | basics| `face_temporal`, `face_spatial`, `unfold`|
| | attributes| `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`|
| |hierarchy|  `switch_key`|
| |matching |   `match_sites`, `match_spatial`, `match_temporal`| 
| |glyph map | `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|

@@

\subsubsection[]{\code{tsibble}}

The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient. A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package. A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument `coords`:

```{r echo = TRUE}
cb <- make_cubble(
  spatial = stations, temporal = meteo_ts, 
  coords = c(long, lat))
```

When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts})  and operations available to the \code{tsibble} class are still valid on these elements. For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:

```{r echo = TRUE}
cb %>% face_temporal() %>% tsibble::has_gaps()
```

\subsubsection[]{\code{sf} (\code{s2})}

*add an example on changing CRS?*

The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). The package \pkg{sf} provides functions that operate efficiently on this spatial information. A \code{cubble} object can store spatial information in  the \code{sf} class. Methods for the \code{sfc} class can be applied in the nested form of the cubble object.  An illustration is in Section \@ref(covid). The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.

```{r echo = TRUE}
cb <- make_cubble(
  spatial = stations_sf, temporal = meteo, 
  key = id, index = date)
class(cb)
#cb %>% sf::st_transform(crs = "EPSG:3857")
```

>>>>

[also talk about the by argument with example]

@@

|Category | Details |  Functions|
|--|---|----------------|
|base R | subset | `[`, `[<-`, `names<-`, `[[`, `[[<-`| 
|tidyverse | dplyr |`dplyr_row_slice`, `dplyr_col_modify`, `dplyr_reconstruct`, `select`, `mutate`, `arrange`, `filter`, `summarise`, `slice`, `group_by`, `ungroup`|
|| tidyr| | 
|| pillar| `type_sum` | 
|sf| | | 
|tsibble| | | 
|cubble | basics| `face_temporal`, `face_spatial`, `unfold`|
| | attributes| `key`, `key_vars`, `key_data`, `index`, `index_var`, `coords`, `spatial`|
| |hierarchy|  `switch_key`|
| |matching |   `match_sites`, `match_spatial`, `match_temporal`| 
| |glyph map | `geom_glyph`, `geom_glyph_box`, `geom_glyph_line`|

@@

Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis. For example, the \code{tbl_ts} class from the \code{tsibble} package [@tsibble] is commonly used in time series forecasting and similarly, the \code{sf} class [@sf] is often used in spatial data science. In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.

### tsibble

The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e. \code{meteo_ts} in the example below. The \code{tsibble} class from the input will be carried over to the \code{cubble} object:

```{r echo = TRUE}
res <- make_cubble(
  spatial = stations, temporal = meteo_ts, coords = c(long, lat))
(res_long <- face_temporal(res))
class(res_long)
```

The long cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}). Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:

```{r echo = TRUE}
res_long %>% tsibble::has_gaps()
```

An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}. The following code illustrates this with the object \code{cb_long} created in section \@ref(create) and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:

```{r echo = TRUE}
cb_long2 <- cb_long %>% make_temporal_tsibble() 
all.equal(cb_long2, res_long)
```

### sf

Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted. This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \@ref(covid)) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons. The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:

```{r echo = TRUE}
(res_nested <- make_cubble(
  spatial = stations_sf, temporal = meteo, 
  key = id, index = date))
class(res_nested)
```

The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \@ref(tab:funs) and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:

```{r echo =TRUE}
#res_nested %>% sf::st_transform(crs = "EPSG:3857")
```

The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:

```{r echo = TRUE}
cb_nested2 <- cb_nested %>% make_spatial_sf() 
all.equal(cb_nested2, res_nested)
```

### 2023-06-01 jss/main.Rmd
combine section 2 and 3; rearrange section 3 with bullet points
rename files
section 2.4 compatibility with tsibble and sf

options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5,
        width = 70, styler.cache_root = "styler-perm")
```

@@

```{r child = here::here("2-cube.Rmd")}

@@

```{r child = here::here("3-the-cubble-pkg.Rmd")}

@@

```{r child = here::here("4-other-features.Rmd")}

@@

```{r child = here::here("5-1-covid.Rmd")}

@@

```{r child = here::here("5-2-aus-hist-max-temp.Rmd")}

@@

```{r child = here::here("5-3-river-level-rainfall.Rmd")}

@@

```{r child = here::here("5-4-era5.Rmd")}

@@

```{r child = here::here("5-5-interactive-graphics.Rmd")}

@@

```{r child = here::here("6-conclusion.Rmd")}

>>>>

options(prompt = "R> ", continue = "+ ",
        tibble.print_max = 5, tibble.print_min = 5,
        width = 90, styler.cache_root = "styler-perm")
```

@@

```{r child = here::here("2-cubble.Rmd")}

@@

```{r child = here::here("3-advanced.Rmd")}

@@

```{r child = here::here("4-1-covid.Rmd")}

@@

```{r child = here::here("4-2-glyph.Rmd")}

@@

```{r child = here::here("4-3-matching.Rmd")}

@@

```{r child = here::here("4-4-netcdf.Rmd")}

@@

```{r child = here::here("4-5-interactive.Rmd")}

@@

```{r child = here::here("5-conclusion.Rmd")}

@@

### 2023-06-01 jss/main.tex
section 2.4 compatibility with tsibble and sf

The \code{key} and \code{index} arguments corresponds to the \code{tsibble} counterpart and \code{coords} derives from the geometry column in an \code{sf} object. Hence, the corresponding argument can be safely omitted, if the spatial data is an \code{sf} object, i.e.~\code{stations_sf}, or the temporal data is a \code{tsibble} object, i.e.~\code{meteo_ts}. The \code{sf} and \code{tsibble} class from the input will be carried over to the \code{cubble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> (res <- make_cubble(spatial = stations_sf, temporal = meteo_ts))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form, [sf]
# extent:   [144.8321, -37.98, 145.0964, -37.6655]
# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
  id           long   lat  elev name  wmo_id            geometry
  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
# i 1 more variable: ts <list>
\end{CodeOutput}
\begin{CodeInput}
R> class(res)
\end{CodeInput}
\begin{CodeOutput}
[1] "spatial_cubble_df" "cubble_df"         "sf"               
[4] "tbl_df"            "tbl"               "data.frame"       
\end{CodeOutput}
\begin{CodeInput}
R> class(res$ts[[1]])
\end{CodeInput}
\begin{CodeOutput}
[1] "tbl_ts"     "tbl_df"     "tbl"        "data.frame"
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7000}}@{}}
\caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Details
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Details
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endhead
base R & subset & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
tidyverse & dplyr & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{slice}, \texttt{group\_by}, \texttt{ungroup} \\
tidyverse & tidyr & \\
tidyverse & pillar & \texttt{type\_sum} \\
sf & & \\
tsibble & & \\
cubble & basics & \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold} \\
& attributes & \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial} \\
& hierarchy & \texttt{switch\_key} \\
& matching & \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal} \\
& glyph map & \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
\bottomrule()
\end{longtable}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_temporal(cb_nested), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_spatial(cb_long), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_temporal(face_spatial(cb_long)), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

@@

\subsubsection[]{\code{tsibble}}

The \code{tsibble} class is a subclass of \code{tibble} where the \code{index} and \code{key} components are used to store temporal and strata information, that makes working with temporal data cognitively efficient. A \code{cubble} object can use the \code{tsibble} class to store the temporal information, and effectively utilize the specialist time series operations in the \pkg{tsibble} package. A \code{tsibble} object can also be casted into a \code{cubble} object through supplying the coordinate information in the argument \texttt{coords}:

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- make_cubble(
+   spatial = stations, temporal = meteo_ts, 
+   coords = c(long, lat))
\end{CodeInput}
\end{CodeChunk}

When a nested \code{cubble} is created, each element in the list-column \code{ts} is in the \code{tsibble} class (labelled \code{tbl_ts}) and operations available to the \code{tsibble} class are still valid on these elements. For example, the code below calculates two time series features (mean and variance) of maximum temperature, utilizing the \code{tsibble} syntax in the \code{cubble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> cb %>% face_temporal() %>% tsibble::has_gaps()
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 3 x 2
  id          .gaps
  <chr>       <lgl>
1 ASN00086038 FALSE
2 ASN00086077 FALSE
3 ASN00086282 FALSE
\end{CodeOutput}
\end{CodeChunk}

\subsubsection[]{\code{sf} (\code{s2})}

\emph{add an example on changing CRS?}

The \code{sf} class is also a subclass of \code{tibble} with a specialized feature geometry list-column (\code{sfc}) to store different geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). The package \pkg{sf} provides functions that operate efficiently on this spatial information. A \code{cubble} object can store spatial information in the \code{sf} class. Methods for the \code{sfc} class can be applied in the nested form of the cubble object. An illustration is in Section \ref{covid}. The spatial information can also be stored as an \code{s2} vector in a \code{cubble} object.

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- make_cubble(
+   spatial = stations_sf, temporal = meteo, 
+   key = id, index = date)
R> class(cb)
\end{CodeInput}
\begin{CodeOutput}
[1] "spatial_cubble_df" "cubble_df"         "sf"               
[4] "tbl_df"            "tbl"               "data.frame"       
\end{CodeOutput}
\begin{CodeInput}
R> #cb %>% sf::st_transform(crs = "EPSG:3857")
\end{CodeInput}
\end{CodeChunk}

>>>>

{[}also talk about the by argument with example{]}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0952}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.7619}}@{}}
\caption{\label{tab:funs} Functions implemented in the \pkg{cubble} package}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Details
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Category
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Details
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} \\
\midrule()
\endhead
base R & subset & \texttt{{[}}, \texttt{{[}\textless{}-}, \texttt{names\textless{}-}, \texttt{{[}{[}}, \texttt{{[}{[}\textless{}-} \\
tidyverse & dplyr & \texttt{dplyr\_row\_slice}, \texttt{dplyr\_col\_modify}, \texttt{dplyr\_reconstruct}, \texttt{select}, \texttt{mutate}, \texttt{arrange}, \texttt{filter}, \texttt{summarise}, \texttt{slice}, \texttt{group\_by}, \texttt{ungroup} \\
& tidyr & \\
& pillar & \texttt{type\_sum} \\
sf & & \\
tsibble & & \\
cubble & basics & \texttt{face\_temporal}, \texttt{face\_spatial}, \texttt{unfold} \\
& attributes & \texttt{key}, \texttt{key\_vars}, \texttt{key\_data}, \texttt{index}, \texttt{index\_var}, \texttt{coords}, \texttt{spatial} \\
& hierarchy & \texttt{switch\_key} \\
& matching & \texttt{match\_sites}, \texttt{match\_spatial}, \texttt{match\_temporal} \\
& glyph map & \texttt{geom\_glyph}, \texttt{geom\_glyph\_box}, \texttt{geom\_glyph\_line} \\
\bottomrule()
\end{longtable}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_temporal(cb_nested), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_spatial(cb_long), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> identical(face_spatial(face_temporal(cb_nested)), cb_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\begin{CodeInput}
R> identical(face_temporal(face_spatial(cb_long)), cb_long)
\end{CodeInput}
\begin{CodeOutput}
[1] FALSE
\end{CodeOutput}
\end{CodeChunk}

@@

Analysts often have their own preferred spatial or temporal data structure, which they may wish to keep for spatio-temporal analysis. For example, the \code{tbl_ts} class from the \code{tsibble} package \citep{tsibble} is commonly used in time series forecasting and similarly, the \code{sf} class \citep{sf} is often used in spatial data science. In \code{cubble}, analysts can combine these two structures together by allowing the spatial component to also be an \code{sf} object and the temporal component to also be a \code{tsibble} object.

\hypertarget{tsibble}{%
\subsubsection{tsibble}\label{tsibble}}

The \code{key} and \code{index} arguments in a \code{cubble} object corresponds to the \code{tsibble} counterparts and they can be safely omitted, if the temporal component is a \code{tsibble} object, i.e.~\code{meteo_ts} in the example below. The \code{tsibble} class from the input will be carried over to the \code{cubble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> res <- make_cubble(
+   spatial = stations, temporal = meteo_ts, coords = c(long, lat))
R> (res_long <- face_temporal(res))
\end{CodeInput}
\begin{CodeOutput}
# cubble:  key: id [3], index: date, long form, [tsibble]
# extent:  2020-01-01 -- 2020-01-10 [1D], no gaps
# spatial: long [dbl], lat [dbl], elev [dbl], name [chr], wmo_id
#   [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00086038 2020-01-01     0  26.8  11  
2 ASN00086038 2020-01-02     0  26.3  12.2
3 ASN00086038 2020-01-03     0  34.5  12.7
4 ASN00086038 2020-01-04     0  29.3  18.8
5 ASN00086038 2020-01-05    18  16.1  12.5
# i 25 more rows
\end{CodeOutput}
\begin{CodeInput}
R> class(res_long)
\end{CodeInput}
\begin{CodeOutput}
[1] "temporal_cubble_df" "cubble_df"          "tbl_ts"            
[4] "tbl_df"             "tbl"                "data.frame"        
\end{CodeOutput}
\end{CodeChunk}

The long cubble shows \code{[tsibble]} in the header to indicate the object also being in a \code{tbl_ts} class (\code{tsibble}). Methods applies to the \code{tbl_ts} class can also be applied to the temporal cubble objects, for example, checking whether the data contain temporal gaps:

\begin{CodeChunk}
\begin{CodeInput}
R> res_long %>% tsibble::has_gaps()
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 3 x 2
  id          .gaps
  <chr>       <lgl>
1 ASN00086038 FALSE
2 ASN00086077 FALSE
3 ASN00086282 FALSE
\end{CodeOutput}
\end{CodeChunk}

An existing \code{cubble} object can promote its temporal component to a \code{tsibble} object by applying \code{make_temporal_tsibble()}. The following code illustrates this with the object \code{cb_long} created in section \ref{create} and the promoted cubble object is equal to the cubble object originally created from a \code{tsibble} object:

\begin{CodeChunk}
\begin{CodeInput}
R> cb_long2 <- cb_long %>% make_temporal_tsibble() 
R> all.equal(cb_long2, res_long)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{sf}{%
\subsubsection{sf}\label{sf}}

Similarly, an \code{sf} object can be supplied as the spatial component to create a \code{cubble} object, with the \code{coords} argument being omitted. This opens up the possibility to represent fixed area with polygons or multipolygons (see Applications \ref{covid}) and the \code{coords} argument will be calculated as the centroids of the (multi)polygons. The \code{[sf]} print in the cubble header suggest an spatial component being also a \code{sf} object:

\begin{CodeChunk}
\begin{CodeInput}
R> (res_nested <- make_cubble(
+   spatial = stations_sf, temporal = meteo, 
+   key = id, index = date))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   key: id [3], index: date, nested form, [sf]
# extent:   [144.8321, -37.98, 145.0964, -37.6655]
# temporal: prcp [dbl], tmax [dbl], tmin [dbl]
  id           long   lat  elev name  wmo_id            geometry
  <chr>       <dbl> <dbl> <dbl> <chr>  <dbl>         <POINT [°]>
1 ASN00086038  145. -37.7  78.4 esse~  95866 (144.9066 -37.7276)
2 ASN00086077  145. -38.0  12.1 moor~  94870   (145.0964 -37.98)
3 ASN00086282  145. -37.7 113.  melb~  94866 (144.8321 -37.6655)
# i 1 more variable: ts <list>
\end{CodeOutput}
\begin{CodeInput}
R> class(res_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] "spatial_cubble_df" "cubble_df"         "sf"               
[4] "rowwise_df"        "tbl_df"            "tbl"              
[7] "data.frame"       
\end{CodeOutput}
\end{CodeChunk}

The \code{sf} functions applicable to a \code{cubble} object have been listed in Table \ref{tab:funs} and the following code shows how to perform coordinate transformation with \code{st_transform} on a cubble object:

\begin{CodeChunk}
\begin{CodeInput}
R> #res_nested %>% sf::st_transform(crs = "EPSG:3857")
\end{CodeInput}
\end{CodeChunk}

The counterpart to promote the spatial component in an existing \code{cubble} to be an \code{sf} object is \code{make_spatial_sf()}:

\begin{CodeChunk}
\begin{CodeInput}
R> cb_nested2 <- cb_nested %>% make_spatial_sf() 
R> all.equal(cb_nested2, res_nested)
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\end{CodeChunk}
